---
title: "Categorical Project"
author: "Jenny Eberling"
date: "11/13/2021"
output: word_document
---

My first step was to convert number of cars from a numeric variable with levels 0,1,2,3,4 to a categorical variable with levels zero, one, two, three +

```{r setup, include=FALSE}
library(tidyverse)
library(VGAM)
library(MASS)
library(lmtest)

bike <- read.csv("C:/Users/eberl/OneDrive/Desktop/bike_buyers_clean.csv")
bike$Cars <- as.factor(bike$Cars)
summary(bike$Cars)
# Make num cars categorical
bike<-bike %>% mutate(Cars = recode(Cars, 
  "0" = 0,                                   
  "1" = 1,
  "2" = 2,
  "3" = 3,
  "4" = 3))
#summary(bike$Cars)

bike$Children <- as.factor(bike$Children)
bike$Gender<-as.factor(bike$Gender)
bike$Marital.Status<-as.factor(bike$Marital.Status)
bike$Education<-as.factor(bike$Education)
bike$Occupation<-as.factor(bike$Occupation)
bike$Home.Owner<-as.factor(bike$Home.Owner)
bike$Commute.Distance<-as.factor(bike$Commute.Distance)
bike$Region<-as.factor(bike$Region)
bike$Purchased.Bike<-as.factor(bike$Purchased.Bike)


#Put in a more reasonable order
bike$Commute.Distance <- factor(bike$Commute.Distance, levels = c("0-1 Miles", "1-2 Miles", "2-5 Miles","5-10 Miles","10+ Miles"))

bike$Education <- factor(bike$Education, levels = c("Partial High School", "High School", "Partial College","Bachelors","Graduate Degree"))
```

Data exploration

```{r }
cor(bike$Income,bike$Age) # Not super correlated
plot(bike$Income, bike$Age)

# Income by categorical variables
plot(bike$Income ~ bike$Gender, xlab="Gender",ylab="Income") # Suspiciously equal
plot(bike$Income ~ bike$Marital.Status) # Pretty even
plot(bike$Income ~ bike$Education, xlab="Education", ylab="Income") #positive trend
plot(bike$Income ~ bike$Home.Owner)
plot(bike$Income ~ bike$Commute.Distance, xlab="Distance",ylab="Income") #positive trend
plot(bike$Income ~ bike$Region, xlab="Region",ylab="Income")
plot(bike$Income ~ bike$Purchased.Bike)
plot(bike$Income ~ bike$Children, xlab="Children",ylab="Income", main="Income vs Children")
plot(bike$Income ~ bike$Occupation, xlab="Occupation", ylab="Income")
aggregate(bike$Income, list(bike$Region), FUN=mean)
aggregate(bike$Income, list(bike$Occupation), FUN=mean)


# Age by categorical variabels
plot(bike$Age ~ bike$Gender)  
plot(bike$Age ~ bike$Marital.Status)  
plot(bike$Age ~ bike$Education)
plot(bike$Age ~ bike$Home.Owner)
plot(bike$Age ~ bike$Commute.Distance, xlab="Distance",ylab="Age") #Bit of a trend, younger people living closer
plot(bike$Age ~ bike$Children)
plot(bike$Age ~ bike$Region)
plot(bike$Age ~ bike$Purchased.Bike) #Typically younger people are buying bikes

# Contingency tables of categorical variables
table(bike$Purchased.Bike, bike$Gender)
table(bike$Purchased.Bike, bike$Education)
table(bike$Gender, bike$Education)
table(bike$Purchased.Bike, bike$Commute.Distance)
table(bike$Gender, bike$Commute.Distance)
table(bike$Region, bike$Commute.Distance)
table(bike$Home.Owner, bike$Commute.Distance)
table(bike$Home.Owner, bike$Gender)
table(bike$Purchased.Bike, bike$Marital.Status)
table(bike$Home.Owner, bike$Region)
table(bike$Region, bike$Education)
table(bike$Education, bike$Occupation)
table(bike$Education, bike$Commute.Distance)
table(bike$Children, bike$Region)

# Look at Cars breakdown
table(bike$Cars, bike$Education)
table(bike$Cars, bike$Commute.Distance)
table(bike$Cars, bike$Occupation)
table(bike$Cars, bike$Region)

#For presentation
plot(bike$Income ~ bike$Region, xlab="Region",ylab="Income",col=c("darkolivegreen1", "darkolivegreen3","darkolivegreen4"))
plot(bike$Income ~ bike$Commute.Distance, xlab="Commute Distance",ylab="Income",col=c("darkolivegreen1", "darkolivegreen2","darkolivegreen3","darkolivegreen4","darkolivegreen"))
plot(bike$Income ~ bike$Gender, xlab="Gender",ylab="Income",col=c("darkolivegreen1", "darkolivegreen3"))

#For paper
levels(bike$Education)
label=c('Partial HS', 'HS','Partial College','Bachelors','Graduate')
plot(bike$Income ~ bike$Education, xlab="Education", ylab="Income", names=label, main='Income by Education Level')
plot(bike$Income ~ bike$Gender, xlab="Gender",ylab="Income", main="Income by Gender")
```
cumulative logit model with proportional odds 

Forward selection manually

```{r selection}
#Step 1 - variables individually 

#Forward Selection

#Marital Status - 2694.7
fit1 <- vglm(bike$Cars ~ bike$Marital.Status, family=cumulative(parallel=T))
AIC(fit1)

#Gender - 2690.103
fit1 <- vglm(bike$Cars ~ bike$Gender, family=cumulative(parallel=T))
AIC(fit1)

#Income - 2514.081, p-value = <2e-16 ***
fit1 <- vglm(bike$Cars ~ bike$Income, family=cumulative(parallel=T))
AIC(fit1)

#Children - 2550.5
fit1 <- vglm(bike$Cars ~ bike$Children, family=cumulative(parallel=T))
AIC(fit1)

#Education - 2531.964
fit1 <- vglm(bike$Cars ~ bike$Education, family=cumulative(parallel=T))
AIC(fit1)

#Occupation - 2545.247
fit1 <- vglm(bike$Cars ~ bike$Occupation, family=cumulative(parallel=T))
AIC(fit1)

#Home.Owner - 2688.315, p-value = 0.0113 *
fit1 <- vglm(bike$Cars ~ bike$Home.Owner, family=cumulative(parallel=T))
AIC(fit1)

#Commute Distance - 2480.624
fit1 <- vglm(bike$Cars ~ bike$Commute.Distance, family=cumulative(parallel=T))
AIC(fit1)

#Region - 2622.393
fit1 <- vglm(bike$Cars ~ bike$Region, family=cumulative(parallel=T))
AIC(fit1)

#Age - 2650.816, p-value = 1.91e-10 ***
fit1 <- vglm(bike$Cars ~ bike$Age, family=cumulative(parallel=T))
AIC(fit1)

#Purchased.Bike - 2648.411, p-value = 1.24e-11 ***
fit1 <- vglm(bike$Cars ~ bike$Purchased.Bike, family=cumulative(parallel=T))
AIC(fit1)

```
Add commute.distance and go to step 2
Starting with AIC = 2480.624

```{r }
#Marital.Status - 2481.954
fit2 <- vglm(Cars ~ Commute.Distance + Marital.Status, data=bike, family=cumulative(parallel=T))
AIC(fit2)

#Gender
fit2 <- vglm(Cars ~ Commute.Distance + Gender, data=bike, family=cumulative(parallel=T))
AIC(fit2)

#Income
fit2 <- vglm(Cars ~ Commute.Distance + Income, data=bike, family=cumulative(parallel=T))
AIC(fit2)

#Children
fit2 <- vglm(Cars ~ Commute.Distance + Children, data=bike, family=cumulative(parallel=T))
AIC(fit2)

#Education
fit2 <- vglm(Cars ~ Commute.Distance + Education, data=bike, family=cumulative(parallel=T))
AIC(fit2)

#Occupation
fit2 <- vglm(Cars ~ Commute.Distance + Occupation, data=bike, family=cumulative(parallel=T))
AIC(fit2)

#Home.Owner
fit2 <- vglm(Cars ~ Commute.Distance + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit2)

#Region
fit2 <- vglm(Cars ~ Commute.Distance + Region, data=bike, family=cumulative(parallel=T))
AIC(fit2)

#Age
fit2 <- vglm(Cars ~ Commute.Distance + Age, data=bike, family=cumulative(parallel=T))
AIC(fit2)

#Purchased.Bike
fit2 <- vglm(Cars ~ Commute.Distance + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit2)


```

Add Education
Current AIC = 2348.766

```{r }
#Marital.Status
fit3 <- vglm(Cars ~ Commute.Distance + Education + Marital.Status, data=bike, family=cumulative(parallel=T))
AIC(fit3)

#Gender
fit3 <- vglm(Cars ~ Commute.Distance + Education + Gender, data=bike, family=cumulative(parallel=T))
AIC(fit3)

#Income
fit3 <- vglm(Cars ~ Commute.Distance + Education + Income, data=bike, family=cumulative(parallel=T))
AIC(fit3)

#Children
fit3 <- vglm(Cars ~ Commute.Distance + Education + Children, data=bike, family=cumulative(parallel=T))
AIC(fit3)

#Occupation - AIC = 2036.45
fit3 <- vglm(Cars ~ Commute.Distance + Education + Occupation, data=bike, family=cumulative(parallel=T))
AIC(fit3)

#Home.Owner
fit3 <- vglm(Cars ~ Commute.Distance + Education + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit3)

#Region
fit3 <- vglm(Cars ~ Commute.Distance + Education + Region, data=bike, family=cumulative(parallel=T))
AIC(fit3)

#Age
fit3 <- vglm(Cars ~ Commute.Distance + Education + Age, data=bike, family=cumulative(parallel=T))
AIC(fit3)

#Purchased.Bike
fit3 <- vglm(Cars ~ Commute.Distance + Education + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit3)
```

Add Occupation 
Current AIC = 2036.45

```{r }
#Marital.Status
fit4 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Marital.Status, data=bike, family=cumulative(parallel=T))
AIC(fit4)

#Gender
fit4 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Gender, data=bike, family=cumulative(parallel=T))
AIC(fit4)

#Income
fit4 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income, data=bike, family=cumulative(parallel=T))
AIC(fit4)

#Children
fit4 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Children, data=bike, family=cumulative(parallel=T))
AIC(fit4)

#Home.Owner
fit4 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit4)

#Region
fit4 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Region, data=bike, family=cumulative(parallel=T))
AIC(fit4)

#Age
fit4 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Age, data=bike, family=cumulative(parallel=T))
AIC(fit4)

#Purchased.Bike
fit4 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit4)

```
Add Income
Current AIC = 1960.879

```{r }
#Marital.Status
fit5 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Marital.Status, data=bike, family=cumulative(parallel=T))
AIC(fit5)

#Gender
fit5 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Gender, data=bike, family=cumulative(parallel=T))
AIC(fit5)

#Children
fit5 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children, data=bike, family=cumulative(parallel=T))
AIC(fit5)

#Home.Owner
fit5 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit5)

#Region
fit5 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Region, data=bike, family=cumulative(parallel=T))
AIC(fit5)

#Age 
fit5 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Age, data=bike, family=cumulative(parallel=T))
AIC(fit5)

#Purchased.Bike
fit5 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit5)

```

Add Children
Current AIC = 1901.528

```{r }
#Marital.Status
fit6 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Marital.Status, data=bike, family=cumulative(parallel=T))
AIC(fit6)

#Gender
fit6 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Gender, data=bike, family=cumulative(parallel=T))
AIC(fit6)

#Home.Owner
fit6 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit6)

#Region
fit6 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region, data=bike, family=cumulative(parallel=T))
AIC(fit6)

#Age
fit6 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Age, data=bike, family=cumulative(parallel=T))
AIC(fit6)

#Purchased.Bike
fit6 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit6)

```
Add Region
Current AIC = 1886.925
```{r }
#Marital.Status
fit7 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Marital.Status, data=bike, family=cumulative(parallel=T))
AIC(fit7)

#Gender
fit7 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Gender, data=bike, family=cumulative(parallel=T))
AIC(fit7)

#Home.Owner
fit7 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit7)

#Age
fit7 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Age, data=bike, family=cumulative(parallel=T))
AIC(fit7)

#Purchased.Bike
fit7 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit7)

```
Add Purchased.Bike - JUST BARELY helps
Current AIC = 1875.33
```{r }
#Marital.Status
fit8 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Marital.Status, data=bike, family=cumulative(parallel=T))
AIC(fit8)

#Gender
fit8 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Gender, data=bike, family=cumulative(parallel=T))
AIC(fit8)

#Home.Owner
fit8 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit8)

#Age
fit8 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Age, data=bike, family=cumulative(parallel=T))
AIC(fit8)

```
Add Home.Owner
Current AIC = 1870.065

```{r }
#Marital.Status
fit9 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Marital.Status, data=bike, family=cumulative(parallel=T))
AIC(fit9)

#Gender
fit9 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Gender, data=bike, family=cumulative(parallel=T))
AIC(fit9)

#Age
fit9 <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Age, data=bike, family=cumulative(parallel=T))
AIC(fit9)

```
Finally, none of those are lower, so we've reached out final model. It contains:
Commute.Distance, Education, Occupation, Income, Children, Region, Purchased.Bike, and Home.Owner

```{r}
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))
summary(fit)
AIC(fit)
```








############################################################################################






I'm going to take a look at how checking to remove variables after each addition changes things.

Forward selection with removal

After adding commute.distance, we added education
Having both still has a lower AIC than either on its own. 

```{r }

#Marital Status
fit1 <- vglm(Cars ~ Commute.Distance, data=bike, family=cumulative(parallel=T))
AIC(fit1)

#Add education
fit2 <- vglm(Cars ~ Commute.Distance + Education, data=bike, family=cumulative(parallel=T))
AIC(fit2)

#Marital Status
fit1 <- vglm(Cars ~ Education, data=bike, family=cumulative(parallel=T))
AIC(fit1)


```
Then we added occupation 
Neither of removed models have a lower AIC, so we keep them all in

```{r }
#All three, AIC = 2036.45
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation, data=bike, family=cumulative(parallel=T))
AIC(fit)

#Remove Commute.Distance
fit2 <- vglm(Cars ~ Education + Occupation, data=bike, family=cumulative(parallel=T))
AIC(fit)

#Remove Education
fit <- vglm(Cars ~ Commute.Distance + Occupation, data=bike, family=cumulative(parallel=T))
AIC(fit)

```
Then we added Income
All four has the lowest AIC so no removal

```{r }
#All four
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income, data=bike, family=cumulative(parallel=T))
AIC(fit)

#Without Commute.Distance
fit <- vglm(Cars ~ Education + Occupation + Income, data=bike, family=cumulative(parallel=T))
AIC(fit)

#Without Education
fit <- vglm(Cars ~ Commute.Distance + Occupation + Income, data=bike, family=cumulative(parallel=T))
AIC(fit)

#Without Occupation
fit <- vglm(Cars ~ Commute.Distance + Education + Income, data=bike, family=cumulative(parallel=T))
AIC(fit)

```

Then we added Children
All five is the lowest AIC so keep them all in

```{r }
#All five
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children, data=bike, family=cumulative(parallel=T))
AIC(fit)

#Without Income
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation  + Children, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Without Occupation
fit <- vglm(Cars ~ Commute.Distance + Education + Income + Children, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Without Education
fit <- vglm(Cars ~ Commute.Distance + Occupation + Income + Children, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Without Commute.Distance
fit <- vglm(Cars ~ Education + Occupation + Income + Children, data=bike, family=cumulative(parallel=T))
AIC(fit)
```
Next we add Region
All six is still the lowest, keep all variables

```{r }
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region, data=bike, family=cumulative(parallel=T))
AIC(fit)

#Remove Children
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Region, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Income
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Children + Region, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Occupation
fit <- vglm(Cars ~ Commute.Distance + Education + Income + Children + Region, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Education
fit <- vglm(Cars ~ Commute.Distance + Occupation + Income + Children + Region, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Commute.Distance
fit <- vglm(Cars ~ Education + Occupation + Income + Children + Region, data=bike, family=cumulative(parallel=T))
AIC(fit)

```
Then we add Purchased.Bike
All seven has the lowest AIC, so don't drop any variables
```{r }
#All seven
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit)

#Remove Region
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Children
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Region + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Income
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Children + Region + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Occupation
fit <- vglm(Cars ~ Commute.Distance + Education + Income + Children + Region + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Education
fit <- vglm(Cars ~ Commute.Distance + Occupation + Income + Children + Region + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Commute.Distance
fit <- vglm(Cars ~ Education + Occupation + Income + Children + Region + Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fit)
```
The last variable added was Home.Owner
All eight has the lowest AIC, so no need to remove any. Our final model is the same. 

```{r }
#All eight
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit)

#Remove Purchased.Bike
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Region
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Children
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Income
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Occupation
fit <- vglm(Cars ~ Commute.Distance + Education + Income + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Education
fit <- vglm(Cars ~ Commute.Distance + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit)
#Remove Commute.Distance
fit <- vglm(Cars ~ Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fit)

```



####################################################################################################


Check for interaction
Current low AIC: 1870.065

First, let's look at interactions I think are most plausible. 

```{r }

#CD:R - AIC = 1802.951
fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Commute.Distance:Region, data=bike, family=cumulative(parallel=T))
AIC(fiti)

#Income:HO
fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Income:Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fiti)

#Education:Income
fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Education:Income, data=bike, family=cumulative(parallel=T))
AIC(fiti)

#Commute.Distance:Purchased.Bike
fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Commute.Distance:Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fiti)

#Occupation:Salary
fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Occupation:Income, data=bike, family=cumulative(parallel=T))
AIC(fiti)




#This don't run for some reason
#Occupation:Education
#fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Occupation:Education, data=bike, family=cumulative(parallel=T))
#AIC(fiti)



```


```{r }
########## Commute.Distance Interaction #############
#smallest: 1802.059

#CD:E
fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Commute.Distance:Education, data=bike, family=cumulative(parallel=T))
AIC(fiti)

#CD:I
fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Commute.Distance:Income, data=bike, family=cumulative(parallel=T))
AIC(fiti)

#CD:R
fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Commute.Distance:Region, data=bike, family=cumulative(parallel=T))
AIC(fiti)
#CD:PB
fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Commute.Distance:Purchased.Bike, data=bike, family=cumulative(parallel=T))
AIC(fiti)
#CD:HO
fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Commute.Distance:Home.Owner, data=bike, family=cumulative(parallel=T))
AIC(fiti)



#Doesn't run for some reason? Maybe not enough observations
#CD:Child
#fitin <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Commute.Distance:Children, data=bike, family=cumulative(parallel=T))
#AIC(fitin)

#CD:Oc
#fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Commute.Distance:Occupation, data=bike, family=cumulative(parallel=T))
#AIC(fiti)

```

I have two interactions tied, I will try to add them together in a model

```{r }
#CD:E
fiti <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner + Commute.Distance:Education + Commute.Distance:Region, data=bike, family=cumulative(parallel=T))
AIC(fiti)

```


But we'll still stick with original final model
```{r }
fit <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))

AIC(fit)
```





What if commute distance was ordinal?

H0: It’s okay to treat commute distance as quantitative with scores 0,1,2,5,10
Ha: we need to treat commute distance as categorical

p= 0.5993 so FTR H0. Not sig evidence we need to treat income as categorical.

When scores are 0,1,2,3,4 , p = 0.02488 so reject H0. There is statistically sig evidence we need to treat c.d. as categorical.

```{r }

bike<-bike %>% mutate(Distance.Ordinal = recode(Commute.Distance, 
  "0-1 Miles" = 0,                                   
  "1-2 Miles" = 1,
  "2-5 Miles" = 2,
  "5-10 Miles" = 5,
  "10+ Miles" = 10))


fitord <- vglm(Cars ~ Distance.Ordinal + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))

VGAM::lrtest(fit,fitord)

AIC(fit)
AIC(fitord) # It does decrease by 5


# With different weights
#### now with different weights
bike<-bike %>% mutate(Distance.Ordinal2 = recode(Commute.Distance, 
  "0-1 Miles" = 0,                                   
  "1-2 Miles" = 1,
  "2-5 Miles" = 2,
  "5-10 Miles" = 3,
  "10+ Miles" = 4))

fitord2 <- vglm(Cars ~ Distance.Ordinal2 + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))

AIC(fit)
AIC(fitord2) # It actually increases
VGAM::lrtest(fit,fitord2)

```

What if education was nominal?

H0: It’s okay to treat education as quantitative with scores 0,1,2,3,4
Ha: we need to treat education as categorical

p = 0.001088 so reject H0. There is significant evidence that we need to treat income as categorical.

```{r }
bike<-bike %>% mutate(Education.Ordinal = recode(Education, 
  "Partial High School" = 0,                                   
  "High School" = 1,
  "Partial College" = 2,
  "Bachelors" = 3,
  "Graduate Degree" = 4))


fitedord <- vglm(Cars ~ Commute.Distance + Education.Ordinal + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T)) #with distance.ordinal, decreases by 5 like it did  before

AIC(fit)
AIC(fitedord) # Increases by 10
VGAM::lrtest(fit,fitedord)
```

What if Children was nominal?

H0: It’s okay to treat children as quantitative with scores 0,1,2,3,4,5
Ha: we need to treat children as categorical

p = 6.318e-07 so reject H0. There is statistically significant evidence that we need to treat children as categorical.

```{r }
bike2<-bike
bike2$ch <- bike2$Children
bike2$cho <- (as.numeric(bike2$ch) - 1)
bike<-bike %>% mutate(Children.Ordinal = as.numeric(Children)-1)


fitchord <- vglm(Cars ~ Commute.Distance + Education + Occupation + Income + Children.Ordinal + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T)) #with distance.ordinal, decreases by 5 like it did  before

AIC(fit)
AIC(fitchord) # Increases by 26
VGAM::lrtest(fit,fitchord)
```


################################################################
Now we have our model. Lets look at it.
################################################################


Check proportional odds assumption - model might be overfitted/too complex

```{r }
fit_ods <- vglm(Cars ~ Distance.Ordinal + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=T))
summary(fit_ods)


fit_no <- vglm(Cars ~ Distance.Ordinal + Education + Occupation + Income + Children + Region + Purchased.Bike + Home.Owner, data=bike, family=cumulative(parallel=F))
summary(fit_no)
```
Look at standardized residuals
Who are the massive outliers? Man who commutes a long distance and has 3 kids but 0 Cars
Manager with 0 cars
All have either 0 or 1 cars, but huge outliers have 0 

```{r }
bike$Residual <- resid(fitord)/sqrt(1-hatvalues(fitord))

outliers <- bike %>%
  filter(Residual > 3 | Residual < -3)


```



Predict for myself

```{r }
-2.337+3.750-.254-.00003471*20000-.3527
-.1016+3.750-.254-.00003471*20000-.3527
2.865+3.750-.254-.00003471*20000-.3527

exp( 0.1121)/(1+exp( 0.1121))
exp(2.3475)/(1+exp(2.3475))
exp(5.3141)/(1+exp(5.3141))

0.9127353-0.5279957
0.9951024-0.9127353
1-0.9951024
```



Predict for myself in many years


```{r }
-2.337-.1294*2+5.967-2.330-.00003471*120000-1.004+.4261
-.1016-.1294*2+5.967-2.330-.00003471*120000-1.004+.4261
2.865-.1294*2+5.967-2.330-.00003471*120000-1.004+.4261

exp(-3.7019)/(1+exp(-3.7019))
exp(-1.4665)/(1+exp(-1.4665))
exp(1.5001)/(1+exp(1.5001))

0.1874752-0.02408233
0.8175894-0.1874752
1-0.8175894
```















#################################################################################
Secondary Modeling
#################################################################################

Now we're going to build a model predicting if someone owns any car vs no car and see how it compares

```{r }
bike<-bike %>% mutate(Own.Car = recode(Cars, 
  "0" = 0,                                   
  "1" = 1,
  "2" = 1,
  "3" = 1,
  "4" = 1))
bike$Own.Car <- as.factor(bike$Own.Car)
summary(bike$Own.Car)


bike<-bike %>% mutate(Distance.Ordinal = recode(Commute.Distance, 
  "0-1 Miles" = 0,                                   
  "1-2 Miles" = 1,
  "2-5 Miles" = 2,
  "5-10 Miles" = 5,
  "10+ Miles" = 10))






fullmodel <- glm(Own.Car ~ Marital.Status + Gender + Income + Children + Education + Occupation + Home.Owner + Commute.Distance + Region + Age + Purchased.Bike, data=bike, family=binomial)


stepAIC(fullmodel, direction="forward")

stepAIC(fullmodel, direction="backward") # 550.5
```

Now with Commute.Distance as ordinal

```{r }

fullmodel2 <- glm(Own.Car ~ Marital.Status + Gender + Income + Children + Education + Occupation + Home.Owner + Distance.Ordinal + Region + Age + Purchased.Bike, data=bike, family=binomial)

stepAIC(fullmodel2, direction="forward") # AIC = 559.1
stepAIC(fullmodel2, direction="backward") # AIC = 551.8



#### now with different weights
bike<-bike %>% mutate(Distance.Ordinal2 = recode(Commute.Distance, 
  "0-1 Miles" = 0,                                   
  "1-2 Miles" = 1,
  "2-5 Miles" = 2,
  "5-10 Miles" = 3,
  "10+ Miles" = 4))

fullmodel3 <- glm(Own.Car ~ Marital.Status + Gender + Income + Children + Education + Occupation + Home.Owner + Distance.Ordinal2 + Region + Age + Purchased.Bike, data=bike, family=binomial)

stepAIC(fullmodel3, direction="backward") # AIC = 545.8

```










Best subset strategy based on AIC
library(bestglm)
VARS <- cbind(#vars here)
bestglm(VARS, IC="AIC", family=binomial)






